---
title: "ggregular_test"
author: "DHJ"
date: "23 4 2022"
output: html_document
---

## Packages  
```{r}

library(ggplot2)
library(dplyr)

```


## Make geom without label     
* Based on 'geom_springs' example    
    - Script 10 in Elveovervåkning (C:\Data\seksjon 318\Elveovervakning)  
    - The original example was based on the superb [flipbook](https://evamaerey.github.io/flipbooks/extending_ggplot2.html#1) by Eva Mae Rey which again is based on [The 'spring' chapters](https://ggplot2-book.org/spring1.html) in the ggplot2 book     

### Define create_linecoordinates  
Building on our experiment above , we define a general function for drawing the line to the label           
* **This function will be used in all parts 3.1-3.5**      
```{r}

create_linecoordinates <- function(x, y, angle, length) {
  if (length <= 0) {
    rlang::abort("`tension` must be larger than 0")
  }
  angle_rad <- angle/180*pi
  length_abs <- length
  # length_abs <- length/100*xrange     # if length in percentage of xrange
  x_abs <- length_abs*cos(angle_rad)
  y_abs <- length_abs*cos(angle_rad)
  xend <- x + x_abs
  yend = y + y_abs

# Create the new data
  data.frame(
    x = c(x, xend),
    y = c(y, yend)
  )
}



```

#### Test of our function    
```{r, fig.width=5, fig.height=3}

create_linecoordinates(
  x = 4, y = 2,
  angle = 45, length = 5
) %>%
  ggplot() +
  aes(x = x, y = y) +
  geom_point() +
  geom_path()

```

### Define the stat  

* Always camel-case name (here, 'StatReglabel'), and the new class is always saved into a variable of the same name (`StatSpring <- ggproto("StatSpring", ...`)      
* We make the function by overwriting existing methods (some of which should not be overwritten)  
    - Use `print(Stat)` to see them all  
* We here overwrite three methods:   
    - setup_params   
    - setup_data    
    - compute_panel   
* We could also have overwritten compute_group and compute_layer   
    
```{r}

StatReglabel <- ggproto(
  
  "StatReglabel",    # name
  
  Stat,            # define this as a stat  
  
  # Sets up parameters correctly
  setup_params = function(data, params) {
    if (is.null(params$angle)) {
      params$diameter <- 45
    }
    if (is.null(params$length)) {
      params$tension <- 0.5
    } else if (params$length < 0) {
      rlang::abort("Length must be zero or a positive number")
    }
    params
  },
  
  # Sets up data correctly - this also seems to be boilerplate  
  # If group doesn't have unique numbers, we make it so  
  setup_data = function(data, params) {
    if (anyDuplicated(data$group)) {
      data$group <- paste(data$group, seq_len(nrow(data)), sep = "-")
    }
    data
  },
  
  # There are three possible compute_ methods. We overwrite only compute_panel
  #>     compute_group: function
  #>     compute_layer: function
  #>     compute_panel: function  
  compute_panel = function(data, scales, angle = 45, length = 0.5) {
    cols_to_keep <- setdiff(names(data), c("x", "y"))
    label_lines <- lapply(
      seq_len(nrow(data)), 
      function(i) {
      line_path <- create_linecoordinates(data$x[i], data$y[i], angle, length)
      cbind(line_path, unclass(data[i, cols_to_keep]))
      }
      )
    do.call(rbind, label_lines)
  },
  
  # Set what is needed input  
  required_aes = c("x", "y")

  )
    
```

#### Test the stat  

Unpacking the compute_panel function. We start by making some data:     
```{r}

my_data <- tibble(
  x = 1:2,
  y = 4:3,) %>%
  mutate(group = row_number())

angle = 45
length = 5


```


This line in StatSpring just returns all variables that are not the start and end coordinates    
```{r}

cols_to_keep <- setdiff(names(my_data), c("x", "y")) 

```



This part, using `lapply`, creates a list of one data frame per row. We apply it to our data:     
```{r}

# Code as in the stat, excep t that data is replaced by my_data

label_lines <- lapply(
  seq_len(nrow(my_data)), 
  function(i) {
    line_path <- create_linecoordinates(my_data$x[i], my_data$y[i], angle, length)
    cbind(line_path, unclass(my_data[i, cols_to_keep]))
  }
)

str(label_lines)

```

This combines the list of data frames to a single data frame, by binding them together by row:      
```{r}

do.call(rbind, label_lines) 

```



### Define the geom        
* Also called the *constructor*   
* Mainly boilerplate code  
* Match the argument order and naming used in the geom constructors provided by ggplot2 itself, since users have an internalized expectation about how they work  

```{r}

geom_reglabel <- function(mapping = NULL, data = NULL, 
                        stat = "reglabel",              # here goes the name of our stat
                        position = "identity", ..., 
                        angle = 45, length = 5,    # here goes our stat's parameters 
                        arrow = NULL, lineend = "butt", linejoin = "round", 
                        na.rm = FALSE, show.legend = NA, inherit.aes = TRUE) {
  layer(
    data = data, 
    mapping = mapping, 
    stat = stat, 
    geom = GeomPath,            # GeomPath: set type of geom used to show the plot    
    position = position, 
    show.legend = show.legend, 
    inherit.aes = inherit.aes, 
    params = list(              # here our parameters are sent on to the layer() function  
      angle = angle,
      length = length,
      arrow = arrow, 
      lineend = lineend, 
      linejoin = linejoin, 
      na.rm = na.rm, 
      ...
    )
  )
}

```


### Test geom_reglabel  

First we make some data (and plot them)    
```{r, fig.width=6, fig.height=4}

set.seed(123)
mydata <- tibble(
  x = runif(5, max = 10),
  y = runif(5, max = 10),
  angle = 45,
  length = 5,
  class = 1:5
)

# Basic
ggplot(mydata) +
  aes(x = x, y = y) +
  geom_point() +
  geom_reglabel()

# Test faceting
ggplot(mydata) +
  aes(x = x, y = y) +
  geom_point() +
  geom_reglabel() +
  facet_wrap(~class)

# Test colors
ggplot(mydata) +
  aes(x = x, y = y, color = factor(class)) +
  geom_point() +
  geom_reglabel() +
  facet_wrap(~class)

```




## Make geom with label     
* Based on 'geom_springs' example    
    - Script 10 in Elveovervåkning (C:\Data\seksjon 318\Elveovervakning)  
    - The original example was based on the superb [flipbook](https://evamaerey.github.io/flipbooks/extending_ggplot2.html#1) by Eva Mae Rey which again is based on [The 'spring' chapters](https://ggplot2-book.org/spring1.html) in the ggplot2 book     


### Define the path stat  

```{r}

StatReglabelpath <- ggproto(
  
  "StatReglabelpath",    # name
  
  Stat,            # define this as a stat  
  
  # Sets up parameters correctly
  setup_params = function(data, params) {
    if (is.null(params$angle)) {
      params$diameter <- 45
    }
    if (is.null(params$length)) {
      params$tension <- 0.5
    } else if (params$length < 0) {
      rlang::abort("Length must be zero or a positive number")
    }
    params
  },
  
  # Sets up data correctly - this also seems to be boilerplate  
  # If group doesn't have unique numbers, we make it so  
  setup_data = function(data, params) {
    if (anyDuplicated(data$group)) {
      data$group <- paste(data$group, seq_len(nrow(data)), sep = "-")
    }
    data
  },
  
  # There are three possible compute_ methods. We overwrite only compute_panel
  #>     compute_group: function
  #>     compute_layer: function
  #>     compute_panel: function  
  compute_panel = function(data, scales, angle = 45, length = 0.5) {
    cols_to_keep <- setdiff(names(data), c("x", "y"))
    label_lines <- lapply(
      seq_len(nrow(data)), 
      function(i) {
      line_path <- create_linecoordinates(data$x[i], data$y[i], angle, length)
      cbind(line_path, unclass(data[i, cols_to_keep]))
      }
      )
    do.call(rbind, label_lines)
  },
  
  # Set what is needed input  
  required_aes = c("x", "y")

  )
    
```


### Define the label stat  

```{r}

StatReglabeltext <- ggproto(
  
  "StatReglabeltext",    # name
  
  Stat,            # define this as a stat  
  
  # Sets up parameters correctly
  setup_params = function(data, params) {
    if (is.null(params$angle)) {
      params$diameter <- 45
    }
    if (is.null(params$length)) {
      params$tension <- 0.5
    } else if (params$length < 0) {
      rlang::abort("Length must be zero or a positive number")
    }
    params
  },
  
  # Sets up data correctly - this also seems to be boilerplate  
  # If group doesn't have unique numbers, we make it so  
  setup_data = function(data, params) {
    if (anyDuplicated(data$group)) {
      data$group <- paste(data$group, seq_len(nrow(data)), sep = "-")
    }
    data
  },
  
  # There are three possible compute_ methods. We overwrite only compute_panel
  #>     compute_group: function
  #>     compute_layer: function
  #>     compute_panel: function  
  compute_panel = function(data, scales, angle = 45, length = 0.5) {
    cols_to_keep <- setdiff(names(data), c("x", "y", "label"))
    label_positions <- lapply(
      seq_len(nrow(data)), 
      function(i) {
      label_pos <- create_linecoordinates(data$x[i], data$y[i], angle, length) %>%
        tail(1) %>%
        mutate(label = data$label[i])
      cbind(label_pos, unclass(data[i, cols_to_keep]))
      }
      )
    do.call(rbind, label_positions)
  },
  
  # Set what is needed input  
  required_aes = c("x", "y", "label")

  )
    
```



#### Test the stat  

Unpacking the compute_panel function. We start by making some data:     
```{r}

my_data <- tibble(
  x = 1:2,
  y = 4:3,
  label = letters[1:2]) %>%
  mutate(group = row_number())

angle = 45
length = 5


```


This line in StatSpring just returns all variables that are not the start and end coordinates    
```{r}

cols_to_keep <- setdiff(names(my_data), c("x", "y" ,"label")) 

```



This part, using `lapply`, creates a list of one data frame per row. We apply it to our data:     
```{r}

# Code as in the stat, excep t that data is replaced by my_data

label_positions <- lapply(
      seq_len(nrow(my_data)), 
      function(i) {
      label_pos <- create_linecoordinates(my_data$x[i], my_data$y[i], angle, length) %>%
        tail(1) %>%
        mutate(label = my_data$label[i])
      cbind(label_pos, unclass(my_data[i, cols_to_keep]))
      }
      )

str(label_positions)

```

This combines the list of data frames to a single data frame, by binding them together by row:      
```{r}

do.call(rbind, label_positions) 

ggplot(my_data, aes(x, y)) +
  geom_point() +
  geom_text(data = do.call(rbind, label_positions), aes(label = label))

```



### Define path geom_spring      

```{r}

geom_reglabelpath <- function(mapping = NULL, data = NULL, 
                        stat = "reglabelpath",              # here goes the name of our stat
                        position = "identity", ..., 
                        angle = 45, length = 5,    # here goes our stat's parameters 
                        arrow = NULL, lineend = "butt", linejoin = "round", 
                        na.rm = FALSE, show.legend = NA, inherit.aes = TRUE) {
  layer(
    data = data, 
    mapping = mapping, 
    stat = stat, 
    geom = GeomPath,            # GeomPath: set type of geom used to show the plot    
    position = position, 
    show.legend = show.legend, 
    inherit.aes = inherit.aes, 
    params = list(              # here our parameters are sent on to the layer() function  
      angle = angle,
      length = length,
      arrow = arrow, 
      lineend = lineend, 
      linejoin = linejoin, 
      na.rm = na.rm, 
      ...
    )
  )
}

```



### Define label geom_spring      

```{r}

geom_reglabeltext <- function(mapping = NULL, data = NULL, 
                        stat = "reglabeltext",              # here goes the name of our stat
                        position = "identity", ..., 
                        angle = 45, length = 5,    # here goes our stat's parameters 
                        na.rm = FALSE, show.legend = NA, inherit.aes = TRUE) {
  layer(
    data = data, 
    mapping = mapping, 
    stat = stat, 
    geom = GeomText,            # GeomText: set type of geom used to show the plot    
    position = position, 
    show.legend = show.legend, 
    inherit.aes = inherit.aes, 
    params = list(              # here our parameters are sent on to the layer() function  
      angle = angle,
      length = length,
      na.rm = na.rm, 
      ...
    )
  )
}
  
```



### Tests 

#### Test geom_reglabelpath   

First we make some data (and plot them)    
```{r, fig.width=6, fig.height=4}

set.seed(123)
mydata <- tibble(
  x = runif(5, max = 10),
  y = runif(5, max = 10),
  label = letters[1:5],
  angle = 45,
  length = 5,
  class = 1:5
)


# Test colors
ggplot(mydata) +
  aes(x = x, y = y, color = factor(class)) +
  geom_point() +
  geom_reglabelpath() 

# Test colors
ggplot(mydata) +
  aes(x = x, y = y, label = label, color = factor(class)) +
  geom_point() +
  geom_reglabeltext(aes(label = label), angle = 45, length = 5, na.rm = TRUE) 

```

